---
title: "LegumeWeight"
output: html_document
params: 
  output_dir: "../output"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Legume Weight
Prepared by Loretta Clancy



###Data Definitions:

'Rhizobes', i.e. bacteria living on plant roots, can increase plant growth. This dataset investigates the effect of different rhizobial strains on legume growth. Variable are as follows:

Block: independent experimental 'block'

Plant: plant species (Ad: Acacia dealbata; Am: Acacia mearnsii; Dm: Daviesia mimosoides; Gc: Glycine clandestina; Hv: Hardenbergia violacea; Oe: Oxylobium ellipticum)

Strain: rhizobial strain (note: values are descriptive strain identifiers, not quantitative measures)

Weight: plant weight at conclusion of experiment.

```{r}
library(tidyverse)
getwd()


legumeData <- read_csv("data/legume_weights_rhizobia_20str.csv")
str(legumeData)

```

Certain columns should be made into factors (as character)

```{r}
legumeData$Strain <- as.character(legumeData$Strain)
str(legumeData)
```

I am not sure of any requirement to 'tidy' this data any further .... will look at this later if necessary

and I recall from discussions that "Block" is of no consequence to this analysis (classroom discussion), so need to remember this.

```{r}
ggplot(legumeData, aes(x=Plant, y=weight, colour=Strain)) +
       geom_point()
```
There are a few observations with missing measurements
```{r}
sum(is.na(legumeData$weight))
```
We will filter them out

```{r}
legumeData <- legumeData %>% 
  filter(!is.na(legumeData$weight))

```

how many Plants, Strains are there
```{r}
legumeData %>% select(Plant) %>% distinct()
legumeData %>% select(Strain) %>% distinct()
```

