---
title: "LegumeWeight"
output: html_document

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prepared by Loretta Clancy


###Data Definitions:

'**Rhizobes**', i.e. bacteria living on plant roots, can increase plant growth. This dataset investigates the effect of different rhizobial strains on legume growth. Variable are as follows:

**Block**: independent experimental 'block'

**Plant**: plant species (Ad: Acacia dealbata; Am: Acacia mearnsii; Dm: Daviesia mimosoides; Gc: Glycine clandestina; Hv: Hardenbergia violacea; Oe: Oxylobium ellipticum)

**Strain**: rhizobial strain (note: values are descriptive strain identifiers, not quantitative measures)

**Weight**: plant weight at conclusion of experiment.

```{r}
library(tidyverse)
getwd()


legumeData <- read_csv("../data/legume_weights_rhizobia_20str.csv")
str(legumeData)

```

Certain columns should be made into factors (as character)

```{r}
legumeData$Strain <- as.character(legumeData$Strain)
str(legumeData)
```

I am not sure of any requirement to 'tidy' this data any further .... will look at this later if necessary

and I recall from discussions that "Block" is of no consequence to this analysis (classroom discussion), so need to remember this.

```{r}
ggplot(legumeData, aes(x=Plant, y=weight, colour=Strain)) +
  geom_point() 
```


There are a few observations with missing measurements
```{r}
sum(is.na(legumeData$weight))
```

We will filter them out
```{r}
legumeData <- legumeData %>% 
  filter(!is.na(legumeData$weight))

```

how many Plants, Strains are there
```{r}
legumeData %>% select(Plant) %>% distinct()
legumeData %>% select(Strain) %>% distinct()
```

The above chart doesn't really tell me much, need to try a different view
```{r}
ggplot(legumeData, aes(x=Strain, y=weight, colour=Plant)) +
  geom_point() 

ggplot(legumeData, aes(x=Strain, y=weight, colour=Strain)) +
  geom_point() +
  facet_grid(~Plant)
    
```

Questions:
Does every plant have every strain
Are certain strains related to the weight

```{r}
data.frame(table(legumeData$Plant))

legumeData %>% 
  group_by(Plant, Strain) %>% 
  summarise(count=n())
```
```{r}
ggplot(legumeData, aes(x=Strain, y=weight)) +
  geom_point() +
  facet_grid(Strain~Plant)
```

The above shows that each Plant does have every Strain present ??











