---
title: "LegumeWeight"
output: html_document:
    keep_md: true
---
<!-- 
To render the output to a different directory, knit it using the console:
rmarkdown::render("scripts/LegumeWeight.Rmd", output_file="../output/LegumeWeight.html") 
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prepared by Loretta Clancy


###Data Definitions:

'**Rhizobes**', i.e. bacteria living on plant roots, can increase plant growth. This dataset investigates the effect of different rhizobial strains on legume growth. Variable are as follows:

**Block**: independent experimental 'block'

**Plant**: plant species (Ad: Acacia dealbata; Am: Acacia mearnsii; Dm: Daviesia mimosoides; Gc: Glycine clandestina; Hv: Hardenbergia violacea; Oe: Oxylobium ellipticum)

**Strain**: rhizobial strain (note: values are descriptive strain identifiers, not quantitative measures)

**Weight**: plant weight at conclusion of experiment.

```{r}
library(tidyverse)

legumeData <- read_csv("../data/legume_weights_rhizobia_20str.csv")
str(legumeData)

```

Certain columns should be made into factors (as character)

```{r}
legumeData$Strain <- as.character(legumeData$Strain)
str(legumeData)
```

These are defined here so that they can be used down below.
Note:  Actual names have a line break in them
```{r}
plantNames <- list('Ad'="Acacia\ndealbata", 
                   'Am'="Acacia\nmearnsii", 
                   'Dm'="Daviesia\nmimosoides",
                   'Gc'="Glycine\nclandestina", 
                   'Hv'="Hardenbergia\nviolacea",
                   'Oe'="Oxylobium\nellipticum"
                   )
plantLabeller <- function(variable,value){
  return(plantNames[value])
}
```


I am not sure of any requirement to 'tidy' this data any further .... will look at this later if necessary

and I recall from discussions that "Block" is of no consequence to this analysis (classroom discussion), so need to remember this.

```{r}
ggplot(legumeData, aes(x=Plant, y=weight, colour=Strain)) +
    geom_point() 
```


There are a few observations with missing measurements
```{r}
sum(is.na(legumeData$weight))
```

We will filter them out
```{r}
legumeData <- legumeData %>% 
    filter(!is.na(legumeData$weight))

```

How many Plants, Strains are there
```{r}
legumeData %>% select(Plant) %>% distinct()
legumeData %>% select(Strain) %>% distinct()
```

The above chart doesn't really tell me much, need to try a different view
```{r}
ggplot(legumeData, aes(x=Strain, y=weight, colour=Plant)) +
    geom_point() 

ggplot(legumeData, aes(x=Strain, y=weight, colour=Strain)) +
    geom_point() +
    facet_grid(~Plant, labeller = plantLabeller)
    
```

---

##Question:  Does every plant have every strain?
```{r}
data.frame(table(legumeData$Plant))

checkData <- legumeData %>% 
    group_by(Plant, Strain) %>% 
    summarise(count=n()) 

ggplot(checkData, aes(x=Plant, y=Strain)) +
    geom_point(aes(colour=count, size=count)) +
    guides(colour='legend')
```

---

`**NOTE: Need to find out how to combine these legends into one **`

---


To view this in a data format:
```{r}
gridCheckData <- checkData %>% 
    spread(key=Strain, value = count) 
gridCheckData
```

The above shows that each Plant does have every Strain present, but the counts of the strains vary.

##Questions:  Does the weight have any correlation with the strain for the plants?

```{r}
ggplot(legumeData, aes(x=Plant, weight=weight), position=position_stack(reverse = TRUE)) +
    coord_flip() +
    geom_bar(aes(fill=Strain), colour="black") +
    theme(legend.position = "top")
```

Now lets have a look at the mean weights for the Strains

```{r}
checkData2 <- legumeData %>% 
    group_by(Plant, Strain) %>% 
    summarise(meanWeight = mean(weight)) 

ggplot(checkData2, aes(x=Plant, y=Strain)) +
    geom_tile(aes(fill=meanWeight)) +
    scale_fill_gradient(low="LightBlue", high="DarkBlue", limits=c(0, 3)) +
    labs(title = "Mean Weights of Strains for each Plant", fill="mean Weight")
```

Is this easier to read than

```{r}
ggplot(checkData2, aes(x=Plant, y=Strain, size=meanWeight)) +
    geom_point() +
    labs(title = "Mean Weights of Strains for each Plant", size="mean Weight")
```

Or another alternative
```{r}


ggplot(checkData2, aes(x=Strain, y=meanWeight, colour=Strain)) +
    geom_point() +
    facet_grid(~Plant, labeller=plantLabeller) +
    labs(title = "Mean Weights of Strains for each Plant", 
         y="mean Weight") +
    theme(axis.text.x = element_blank())
```




